---
title: 'GTEx meta-topic model analysis: K=20'
author: "Kushal K Dey"
date: "April 18, 2016"
output: 
  html_document:
    css: floating-toc.css
    toc: true
---

## Overview

We perform GTEx V6 metatopic model analysis for K=20. We do it for three scenarios. 

In the first scenario, we generate meta-topics over $5$ levels. We take $20$ topics at the bottom-most level, $15$ in second, $10$ in third, $5$ in fourth and $2$ in the last.

For the second scenario, we generate meta-topics over $4$ levels. We take $20$ topics at the bottom-most level, $10$ in second, $6$ in third and $3$ in fourth. 

For the third scenario, we consider meta-topics over $3$ levels. We take $20$ topics at the bottom-most level, $5$ topics at the second and $2$ in the third.

For the final scenario, we consider meta-topics over $2$ levels. We take $20$ topics at the bottom level and $2$ levels at the top.


The two-fold goal here is to see which patterns are consistent across the different scenarios. Whether we get organ-related information from any of the metatopic structure. Also, which two tissues separate out at the upper-most metatopic level. Is that consistent across the different scenarios?

Note that this analysis is just a naive way of replicating the essence of Deep Poisson factor models and so, the results might be slightly different from the application of DPFM on this data. But the main reason for applying this is to figure out whether there is any merit to using deep model for clustering RNA-seq data at all. Plus we want something that is closer to a clustering model or a topic model than a factor model and more specifically, something that can be seen as an extension of *maptpx*.

```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
library(data.table)
data <- data.frame(fread('../external_data/GTEX_V6/cis_gene_expression.txt'));
matdata <- data[,-(1:2)];
sample_labels <- read.table("../external_data/GTEX_V6/samples_id.txt",
                            header = TRUE, sep = " ",
                            stringsAsFactors = FALSE)
```

## Scenario 1

```{r echo=TRUE, eval=TRUE}
gtex_20 <- get(load("../rdas/gtexv6fit.k.20.part1.rda"));
gtex_20_omega <- gtex_20$omega
gtex_20_theta <- gtex_20$theta
```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
omega <- gtex_20_omega
colnames(omega) <- c(1:NCOL(omega))

# make cell sample labels
# want a version consistent with majority of the literature
sample_labels <- read.table("../external_data/GTEX_V6/samples_id.txt",
                            header = TRUE, sep = " ",
                            stringsAsFactors = FALSE)
tissue_labels <- vector("numeric", NROW(sample_labels))
tissue_labels <- sample_labels[ ,3]

# clean labels
tissue_labels[grep("Nucleus", tissue_labels)] <- "Brain -N. accumbens"
tissue_labels[grep("Putamen", tissue_labels)] <- "Brain -Putamen"
tissue_labels[grep("Caudate", tissue_labels)] <- "Brain -Caudate"
tissue_labels[grep("Gastroe", tissue_labels)] <- "Esophagus -Gastroesophageal Jn."
tissue_labels[grep("cingulate", tissue_labels)] <- "Brain - Anterior cortex (BA24)."
tissue_labels[grep("EBV", tissue_labels)] <- "Cells -EBV-lymphocytes"
tissue_labels[grep("Suprapubic", tissue_labels)] <- "Skin - Unexposed (Suprapubic)"
tissue_labels[grep("Lower Leg", tissue_labels)] <- "Skin - Sun Exposed (Lower Leg)"

# find sample orders in hierarchical clustering
docweights_per_tissue_mean <- apply(omega, 2,
                                    function(x) { tapply(x, tissue_labels, mean) })
ordering <- heatmap(docweights_per_tissue_mean)$rowInd

# order tissue by hierarhical clustering results
tissue_levels_reordered <- unique(tissue_labels)[ordering]


annotation <- data.frame(
    sample_id = paste0("X", 1:length(tissue_labels)),
    tissue_label = factor(tissue_labels,
                          levels = rev(tissue_levels_reordered ) ) )


cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
          RColorBrewer::brewer.pal(12, "Set3")[c(1,2,5,8,9)],
          RColorBrewer::brewer.pal(9, "Set1")[c(9,7)],
          RColorBrewer::brewer.pal(8, "Dark2")[c(3,4,8)])

cols1 <- sample(cols1, 20, replace=FALSE)

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .1,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 5,
                                 axis_label_face="bold"))

```



### metatopic for K=20

```{r echo=TRUE, eval=FALSE}
libsize_gtex <- colSums(matdata);

omega_1 <- gtex_20_omega
theta_1 <- gtex_20_theta

lambda_1 <- sweep(omega_1, 1, libsize_gtex, "*");

z_level_1 <- array(0,c(dim(lambda_1)[1], dim(theta_1)[1], dim(lambda_1)[2]));
for(k in 1:dim(theta_1)[2]){
  z_level_1[,,k] <- floor(lambda_1[,k]%*%t(theta_1[,k]))
}

z_level_1_counts <- apply(z_level_1, c(1,3), sum);

save(z_level_1_counts, file="../rdas/gtex_20_level_1_data_scene_1.rda")
```

### metatopic for K=15

```{r echo=TRUE, eval=FALSE}
z_level_1_counts <- get(load("../rdas/gtex_20_level_1_data_scene_1.rda"))
topic_clus <- maptpx::topics(z_level_1_counts, K=15, tol=0.1);
save(topic_clus, file="../rdas/gtexv6fit.k.20.part1.scene1.level1.rda")
```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
topic_clus <- get(load("../rdas/gtexv6fit.k.20.part1.scene1.level1.rda"))
omega <- topic_clus$omega
# define colors of the clusers
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
           RColorBrewer::brewer.pal(12, "Set3"))

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .4,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 3,
                                 axis_label_face = "bold"))


```

### metatopic K=10

```{r echo=TRUE, eval=FALSE}
libsize_gtex <- rowSums(z_level_1_counts);

omega_2 <- topic_clus$omega
theta_2 <- topic_clus$theta

lambda_2 <- sweep(omega_2, 1, libsize_gtex, "*");

z_level_2 <- array(0,c(dim(lambda_2)[1], dim(theta_2)[1], dim(lambda_2)[2]));
for(k in 1:dim(theta_2)[2]){
  z_level_2[,,k] <- floor(lambda_2[,k]%*%t(theta_2[,k]))
}

z_level_2_counts <- apply(z_level_2, c(1,3), sum);

```

We apply topic model with $K=10$ on this matrix.

```{r echo=TRUE, eval=FALSE}
topic_clus <- maptpx::topics(z_level_2_counts, K=10, tol=0.1);
save(topic_clus, file="../rdas/gtexv6fit.k.20.part1.scene1.level2.rda")

```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
topic_clus <- get(load("../rdas/gtexv6fit.k.20.part1.scene1.level2.rda"))
omega <- topic_clus$omega
# define colors of the clusers
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
           RColorBrewer::brewer.pal(12, "Set3"))

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .4,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 3,
                                 axis_label_face = "bold"))


```

### metatopic K=5

```{r echo=TRUE, eval=FALSE}
libsize_gtex <- rowSums(z_level_2_counts);

omega_3 <- topic_clus$omega
theta_3 <- topic_clus$theta

lambda_3 <- sweep(omega_3, 1, libsize_gtex, "*");

z_level_3 <- array(0,c(dim(lambda_3)[1], dim(theta_3)[1], dim(lambda_3)[2]));
for(k in 1:dim(theta_3)[2]){
  z_level_3[,,k] <- floor(lambda_3[,k]%*%t(theta_3[,k]))
}

z_level_3_counts <- apply(z_level_3, c(1,3), sum);

```

We apply topic model with $K=5$ on this matrix.

```{r echo=TRUE, eval=FALSE}
topic_clus <- maptpx::topics(z_level_3_counts, K=5, tol=0.1);
save(topic_clus, file="../rdas/gtexv6fit.k.20.part1.scene1.level3.rda")
```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
topic_clus <- get(load(file="../rdas/gtexv6fit.k.20.part1.scene1.level3.rda"))
omega <- topic_clus$omega
# define colors of the clusers
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
           RColorBrewer::brewer.pal(12, "Set3"))

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .4,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 3,
                                 axis_label_face = "bold"))


```

### metatopic K=2

```{r echo=TRUE, eval=FALSE}
libsize_gtex <- rowSums(z_level_3_counts);

omega_4 <- topic_clus$omega
theta_4 <- topic_clus$theta

lambda_4 <- sweep(omega_4, 1, libsize_gtex, "*");

z_level_4 <- array(0,c(dim(lambda_4)[1], dim(theta_4)[1], dim(lambda_4)[2]));
for(k in 1:dim(theta_4)[2]){
  z_level_4[,,k] <- floor(lambda_4[,k]%*%t(theta_4[,k]))
}

z_level_4_counts <- apply(z_level_4, c(1,3), sum);

```

We apply topic model with $K=2$ on this matrix.

```{r echo=TRUE, eval=FALSE}
topic_clus <- maptpx::topics(z_level_4_counts, K=2, tol=0.1);
save(topic_clus, file="../rdas/gtexv6fit.k.20.part1.scene1.level4.rda")
```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
topic_clus <- get(load(file="../rdas/gtexv6fit.k.20.part1.scene1.level4.rda"))
omega <- topic_clus$omega
# define colors of the clusers
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
           RColorBrewer::brewer.pal(12, "Set3"))

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .4,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 3,
                                 axis_label_face = "bold"))


```


## Scenario 2

```{r echo=TRUE, eval=FALSE}
gtex_20 <- get(load("../rdas/gtexv6fit.k.20.part1.rda"));
gtex_20_omega <- gtex_20$omega
gtex_20_theta <- gtex_20$theta
```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
omega <- gtex_20_omega
colnames(omega) <- c(1:NCOL(omega))

# make cell sample labels
# want a version consistent with majority of the literature
sample_labels <- read.table("../external_data/GTEX_V6/samples_id.txt",
                            header = TRUE, sep = " ",
                            stringsAsFactors = FALSE)
tissue_labels <- vector("numeric", NROW(sample_labels))
tissue_labels <- sample_labels[ ,3]

# clean labels
tissue_labels[grep("Nucleus", tissue_labels)] <- "Brain -N. accumbens"
tissue_labels[grep("Putamen", tissue_labels)] <- "Brain -Putamen"
tissue_labels[grep("Caudate", tissue_labels)] <- "Brain -Caudate"
tissue_labels[grep("Gastroe", tissue_labels)] <- "Esophagus -Gastroesophageal Jn."
tissue_labels[grep("cingulate", tissue_labels)] <- "Brain - Anterior cortex (BA24)."
tissue_labels[grep("EBV", tissue_labels)] <- "Cells -EBV-lymphocytes"
tissue_labels[grep("Suprapubic", tissue_labels)] <- "Skin - Unexposed (Suprapubic)"
tissue_labels[grep("Lower Leg", tissue_labels)] <- "Skin - Sun Exposed (Lower Leg)"

# find sample orders in hierarchical clustering
docweights_per_tissue_mean <- apply(omega, 2,
                                    function(x) { tapply(x, tissue_labels, mean) })
ordering <- heatmap(docweights_per_tissue_mean)$rowInd

# order tissue by hierarhical clustering results
tissue_levels_reordered <- unique(tissue_labels)[ordering]


annotation <- data.frame(
    sample_id = paste0("X", 1:length(tissue_labels)),
    tissue_label = factor(tissue_labels,
                          levels = rev(tissue_levels_reordered ) ) )


cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
          RColorBrewer::brewer.pal(12, "Set3")[c(1,2,5,8,9)],
          RColorBrewer::brewer.pal(9, "Set1")[c(9,7)],
          RColorBrewer::brewer.pal(8, "Dark2")[c(3,4,8)])

cols1 <- sample(cols1, 20, replace=FALSE)

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .1,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 5,
                                 axis_label_face="bold"))

```



### metatopic for K=20

```{r echo=TRUE, eval=FALSE}
libsize_gtex <- colSums(matdata);

omega_1 <- gtex_20_omega
theta_1 <- gtex_20_theta

lambda_1 <- sweep(omega_1, 1, libsize_gtex, "*");

z_level_1 <- array(0,c(dim(lambda_1)[1], dim(theta_1)[1], dim(lambda_1)[2]));
for(k in 1:dim(theta_1)[2]){
  z_level_1[,,k] <- floor(lambda_1[,k]%*%t(theta_1[,k]))
}

z_level_1_counts <- apply(z_level_1, c(1,3), sum);

save(z_level_1_counts, file="../rdas/gtex_20_level_1_data_scene_2.rda")
```

### metatopic for K=10

```{r echo=TRUE, eval=FALSE}
z_level_1_counts <- get(load("../rdas/gtex_20_level_1_data_scene_2.rda"))
topic_clus <- maptpx::topics(z_level_1_counts, K=10, tol=0.1);
save(topic_clus, file="../rdas/gtexv6fit.k.20.part1.scene2.level1.rda")
```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
topic_clus <- get(load("../rdas/gtexv6fit.k.20.part1.scene2.level1.rda"))
omega <- topic_clus$omega
# define colors of the clusers
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
           RColorBrewer::brewer.pal(12, "Set3"))

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .4,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 3,
                                 axis_label_face = "bold"))


```

### metatopic K=6

```{r echo=TRUE, eval=FALSE}
libsize_gtex <- rowSums(z_level_1_counts);

omega_2 <- topic_clus$omega
theta_2 <- topic_clus$theta

lambda_2 <- sweep(omega_2, 1, libsize_gtex, "*");

z_level_2 <- array(0,c(dim(lambda_2)[1], dim(theta_2)[1], dim(lambda_2)[2]));
for(k in 1:dim(theta_2)[2]){
  z_level_2[,,k] <- floor(lambda_2[,k]%*%t(theta_2[,k]))
}

z_level_2_counts <- apply(z_level_2, c(1,3), sum);

```

We apply topic model with $K=6$ on this matrix.

```{r echo=TRUE, eval=FALSE}
topic_clus <- maptpx::topics(z_level_2_counts, K=6, tol=0.1);
save(topic_clus, file="../rdas/gtexv6fit.k.20.part1.scene2.level2.rda")

```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
topic_clus <- get(load("../rdas/gtexv6fit.k.20.part1.scene2.level2.rda"))
omega <- topic_clus$omega
# define colors of the clusers
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
           RColorBrewer::brewer.pal(12, "Set3"))

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .4,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 3,
                                 axis_label_face = "bold"))


```

### metatopic K=3

```{r echo=TRUE, eval=FALSE}
libsize_gtex <- rowSums(z_level_2_counts);

omega_3 <- topic_clus$omega
theta_3 <- topic_clus$theta

lambda_3 <- sweep(omega_3, 1, libsize_gtex, "*");

z_level_3 <- array(0,c(dim(lambda_3)[1], dim(theta_3)[1], dim(lambda_3)[2]));
for(k in 1:dim(theta_3)[2]){
  z_level_3[,,k] <- floor(lambda_3[,k]%*%t(theta_3[,k]))
}

z_level_3_counts <- apply(z_level_3, c(1,3), sum);

```

We apply topic model with $K=3$ on this matrix.

```{r echo=TRUE, eval=FALSE}
topic_clus <- maptpx::topics(z_level_3_counts, K=3, tol=0.1);
save(topic_clus, file="../rdas/gtexv6fit.k.20.part1.scene2.level3.rda")

```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
topic_clus <- get(load(file="../rdas/gtexv6fit.k.20.part1.scene2.level3.rda"))
omega <- topic_clus$omega
# define colors of the clusers
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
           RColorBrewer::brewer.pal(12, "Set3"))

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .4,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 3,
                                 axis_label_face = "bold"))


```

## Scenario 3

```{r echo=TRUE, eval=FALSE}
gtex_20 <- get(load("../rdas/gtexv6fit.k.20.part1.rda"));
gtex_20_omega <- gtex_20$omega
gtex_20_theta <- gtex_20$theta
```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
omega <- gtex_20_omega
colnames(omega) <- c(1:NCOL(omega))

# make cell sample labels
# want a version consistent with majority of the literature
sample_labels <- read.table("../external_data/GTEX_V6/samples_id.txt",
                            header = TRUE, sep = " ",
                            stringsAsFactors = FALSE)
tissue_labels <- vector("numeric", NROW(sample_labels))
tissue_labels <- sample_labels[ ,3]

# clean labels
tissue_labels[grep("Nucleus", tissue_labels)] <- "Brain -N. accumbens"
tissue_labels[grep("Putamen", tissue_labels)] <- "Brain -Putamen"
tissue_labels[grep("Caudate", tissue_labels)] <- "Brain -Caudate"
tissue_labels[grep("Gastroe", tissue_labels)] <- "Esophagus -Gastroesophageal Jn."
tissue_labels[grep("cingulate", tissue_labels)] <- "Brain - Anterior cortex (BA24)."
tissue_labels[grep("EBV", tissue_labels)] <- "Cells -EBV-lymphocytes"
tissue_labels[grep("Suprapubic", tissue_labels)] <- "Skin - Unexposed (Suprapubic)"
tissue_labels[grep("Lower Leg", tissue_labels)] <- "Skin - Sun Exposed (Lower Leg)"

# find sample orders in hierarchical clustering
docweights_per_tissue_mean <- apply(omega, 2,
                                    function(x) { tapply(x, tissue_labels, mean) })
ordering <- heatmap(docweights_per_tissue_mean)$rowInd

# order tissue by hierarhical clustering results
tissue_levels_reordered <- unique(tissue_labels)[ordering]


annotation <- data.frame(
    sample_id = paste0("X", 1:length(tissue_labels)),
    tissue_label = factor(tissue_labels,
                          levels = rev(tissue_levels_reordered ) ) )


cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
          RColorBrewer::brewer.pal(12, "Set3")[c(1,2,5,8,9)],
          RColorBrewer::brewer.pal(9, "Set1")[c(9,7)],
          RColorBrewer::brewer.pal(8, "Dark2")[c(3,4,8)])

cols1 <- sample(cols1, 20, replace=FALSE)

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .1,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 5,
                                 axis_label_face="bold"))

```



### metatopic for K=20

```{r echo=TRUE, eval=FALSE}
libsize_gtex <- colSums(matdata);

omega_1 <- gtex_20_omega
theta_1 <- gtex_20_theta

lambda_1 <- sweep(omega_1, 1, libsize_gtex, "*");

z_level_1 <- array(0,c(dim(lambda_1)[1], dim(theta_1)[1], dim(lambda_1)[2]));
for(k in 1:dim(theta_1)[2]){
  z_level_1[,,k] <- floor(lambda_1[,k]%*%t(theta_1[,k]))
}

z_level_1_counts <- apply(z_level_1, c(1,3), sum);

save(z_level_1_counts, file="../rdas/gtex_20_level_1_data_scene_2.rda")
```

### metatopic for K=5

```{r echo=TRUE, eval=FALSE}
z_level_1_counts <- get(load("../rdas/gtex_20_level_1_data_scene_2.rda"))
topic_clus <- maptpx::topics(z_level_1_counts, K=5, tol=0.1);
save(topic_clus, file="../rdas/gtexv6fit.k.20.part1.scene3.level1.rda")
```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
topic_clus <- get(load("../rdas/gtexv6fit.k.20.part1.scene3.level1.rda"))
omega <- topic_clus$omega
# define colors of the clusers
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
           RColorBrewer::brewer.pal(12, "Set3"))

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .4,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 3,
                                 axis_label_face = "bold"))


```

### metatopic K=2

```{r echo=TRUE, eval=FALSE}
libsize_gtex <- rowSums(z_level_1_counts);

omega_2 <- topic_clus$omega
theta_2 <- topic_clus$theta

lambda_2 <- sweep(omega_2, 1, libsize_gtex, "*");

z_level_2 <- array(0,c(dim(lambda_2)[1], dim(theta_2)[1], dim(lambda_2)[2]));
for(k in 1:dim(theta_2)[2]){
  z_level_2[,,k] <- floor(lambda_2[,k]%*%t(theta_2[,k]))
}

z_level_2_counts <- apply(z_level_2, c(1,3), sum);

```

We apply topic model with $K=2$ on this matrix.

```{r echo=TRUE, eval=FALSE}
topic_clus <- maptpx::topics(z_level_2_counts, K=2, tol=0.1);
save(topic_clus, file="../rdas/gtexv6fit.k.20.part1.scene3.level2.rda")

```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
topic_clus <- get(load("../rdas/gtexv6fit.k.20.part1.scene3.level2.rda"))
omega <- topic_clus$omega
# define colors of the clusers
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
           RColorBrewer::brewer.pal(12, "Set3"))

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .4,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 3,
                                 axis_label_face = "bold"))


```


## Scenario 4

```{r echo=TRUE, eval=FALSE}
gtex_20 <- get(load("../rdas/gtexv6fit.k.20.part1.rda"));
gtex_20_omega <- gtex_20$omega
gtex_20_theta <- gtex_20$theta
```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
omega <- gtex_20_omega
colnames(omega) <- c(1:NCOL(omega))

# make cell sample labels
# want a version consistent with majority of the literature
sample_labels <- read.table("../external_data/GTEX_V6/samples_id.txt",
                            header = TRUE, sep = " ",
                            stringsAsFactors = FALSE)
tissue_labels <- vector("numeric", NROW(sample_labels))
tissue_labels <- sample_labels[ ,3]

# clean labels
tissue_labels[grep("Nucleus", tissue_labels)] <- "Brain -N. accumbens"
tissue_labels[grep("Putamen", tissue_labels)] <- "Brain -Putamen"
tissue_labels[grep("Caudate", tissue_labels)] <- "Brain -Caudate"
tissue_labels[grep("Gastroe", tissue_labels)] <- "Esophagus -Gastroesophageal Jn."
tissue_labels[grep("cingulate", tissue_labels)] <- "Brain - Anterior cortex (BA24)."
tissue_labels[grep("EBV", tissue_labels)] <- "Cells -EBV-lymphocytes"
tissue_labels[grep("Suprapubic", tissue_labels)] <- "Skin - Unexposed (Suprapubic)"
tissue_labels[grep("Lower Leg", tissue_labels)] <- "Skin - Sun Exposed (Lower Leg)"

# find sample orders in hierarchical clustering
docweights_per_tissue_mean <- apply(omega, 2,
                                    function(x) { tapply(x, tissue_labels, mean) })
ordering <- heatmap(docweights_per_tissue_mean)$rowInd

# order tissue by hierarhical clustering results
tissue_levels_reordered <- unique(tissue_labels)[ordering]


annotation <- data.frame(
    sample_id = paste0("X", 1:length(tissue_labels)),
    tissue_label = factor(tissue_labels,
                          levels = rev(tissue_levels_reordered ) ) )


cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
          RColorBrewer::brewer.pal(12, "Set3")[c(1,2,5,8,9)],
          RColorBrewer::brewer.pal(9, "Set1")[c(9,7)],
          RColorBrewer::brewer.pal(8, "Dark2")[c(3,4,8)])

cols1 <- sample(cols1, 20, replace=FALSE)

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .1,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 5,
                                 axis_label_face="bold"))

```

### metatopic for K=20

```{r echo=TRUE, eval=FALSE}
libsize_gtex <- colSums(matdata);

omega_1 <- gtex_20_omega
theta_1 <- gtex_20_theta

lambda_1 <- sweep(omega_1, 1, libsize_gtex, "*");

z_level_1 <- array(0,c(dim(lambda_1)[1], dim(theta_1)[1], dim(lambda_1)[2]));
for(k in 1:dim(theta_1)[2]){
  z_level_1[,,k] <- floor(lambda_1[,k]%*%t(theta_1[,k]))
}

z_level_1_counts <- apply(z_level_1, c(1,3), sum);

save(z_level_1_counts, file="../rdas/gtex_20_level_1_data_scene_2.rda")
```

### metatopic for K=2

```{r echo=TRUE, eval=FALSE}
z_level_1_counts <- get(load("../rdas/gtex_20_level_1_data_scene_2.rda"))
topic_clus <- maptpx::topics(z_level_1_counts, K=2, tol=0.1);
save(topic_clus, file="../rdas/gtexv6fit.k.20.part1.scene4.level1.rda")
```

```{r echo=TRUE, eval=TRUE, fig.height=10, fig.width=5, message=FALSE, warning=FALSE}
topic_clus <- get(load("../rdas/gtexv6fit.k.20.part1.scene4.level1.rda"))
omega <- topic_clus$omega
# define colors of the clusers
cols1 <- c(rev(RColorBrewer::brewer.pal(12, "Paired"))[c(3,4,7,8,11,12,5,6,9,10)],
           RColorBrewer::brewer.pal(12, "Set3"))

CountClust::StructureGGplot(omega = omega,
                annotation= annotation,
                palette = cols1,
                yaxis_label = "",
                order_sample = TRUE,
                split_line = list(split_lwd = .4,
                                  split_col = "white"),
                axis_tick = list(axis_ticks_length = .1,
                                 axis_ticks_lwd_y = .1,
                                 axis_ticks_lwd_x = .1,
                                 axis_label_size = 3,
                                 axis_label_face = "bold"))


```
